Lesson 4
========================================================

### Scatterplots
Notes:
Scatterplots used to analyze relationship between two continuous variables
use qplot and pass x and y variables

```{r Scatterplots}
setwd('/Users/danilitovsky/Dropbox/dataScience/R')
library(ggplot2)

facebook = read.csv('pseudo_facebook.tsv', sep = '\t')

qplot(x=age, y=friend_count, data= facebook)
```

***

#### What are some things that you notice right away?
Response:

lots of young people have lots of friends, vertical lines

### ggplot Syntax
Notes:
Start using ggplot instead to specify complicated plots
geom_point() for scatterplots
alpha = 1/20 means that one full point point will represent 20 points, easier to clean data
also used coord_trans to change y ais


```{r ggplot Syntax}
summary(facebook$age)

ggplot(aes(x=age, y = friend_count), data=facebook) + geom_point(alpha = 1/20) + xlim(13,90)

#geom_jitter to replace geom_point to "fill-in" gaps between ages by adding noise. Because it doesn't make visual sense that the ages are all line up virtically. But go back to point when adding coord_trans

ggplot(aes(x=age, y = friend_count), data=facebook) + 
  geom_point(alpha = 1/20) +
  xlim(13,90) + 
  coord_trans(y = "sqrt")

```

#### What do you notice?

***

### Alpha and Jitter
Notes:
geom_jitter ... jitters data, but didn't use it here because it gave an error
why: because jitter makes zero values less than zero, and the squre root of that can be imaginary, so error. To account for this, add position height at least zero
```{r Alpha and Jitter}

ggplot(aes(x=age, y = friend_count), data=facebook) + 
  geom_point(alpha = 1/20, position= position_jitter(h = 0), color = "orange") +
  xlim(13,90) + 
  coord_trans(y = "sqrt")

```


###explore relationship between friendships_initiated and age using the above techniques. Build layer by layer

```{r}
ggplot(aes(x=age, y = friendships_initiated), data=facebook) + geom_point()

ggplot(aes(x=age, y = friendships_initiated), data=facebook) + geom_point() + xlim(13,90)

ggplot(aes(x=age, y = friendships_initiated), data=facebook) + geom_point(color='blue', alpha= 1/20) + xlim(13,90) 
  
#investigate male v. female friendship initiations, add aes to geom)()

ggplot(aes(x=age, y = friendships_initiated), data=facebook) + geom_jitter(aes(color= gender, alpha= 1/20)) + xlim(13,90)  
```

***

### Conditional Means
Notes:
instead of looking at raw values, let's look at the average friends initiated by age

```{r Conditional Means}
#group my data by age
library("dplyr")
age_groups <- group_by(facebook, age)
#summarize the new df and create new variales of means
facebook.fc_by_age <- summarize(age_groups,
          n= n(), #function for summarize that reports number of rows per age group
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count))

# fc_by_age is now a new df telling us the number of people in each age group, and each age group's mean and median friend count
```

#may need to resave new df and order it by age df <- arrange(df, age)

Create your plot using facebook.fc_by_age df mean friend count v. age
and use previous tactics of ggplot

```{r Conditional Means Plot}

ggplot(aes(x=age, y = friend_count_mean), data = facebook.fc_by_age) +
  geom_line(aes(color = "red")) +
  xlim(13,90) +
  coord_trans(y= "log10")

```

***

### Overlaying Summaries with Raw Data
Notes: Use original raw data and overlay above average friend count by using geom_line (stat, fun.y = mean) which will give the mean friend count by age(same we got above)

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x=age, y = friend_count), data=facebook) + 
  geom_point(alpha = 1/20, position= position_jitter(h = 0), color = "orange") +
  xlim(13,90) + 
  coord_trans(y = "sqrt") +
  geom_line(stat = "summary", fun.y = mean)

#use the same plot but now let's add 10, 50, and 90 % quantiles

ggplot(aes(x=age, y = friend_count), data=facebook) + 
  #xlim(13,90) +
  geom_point(alpha = 1/20, position= position_jitter(h = 0), color = "orange") +
  coord_cartesian(xlim = c(13, 70), ylim= c(0, 1000)) + #better than xlim bc data isn't taken out, but take out coord_trans to work
  #coord_trans(y = "sqrt") +
  geom_line(stat = "summary", fun.y = mean, color= "black") +
  geom_line(stat = "summary", fun.y= quantile, probs = .1, color = "blue") +
  geom_line(stat = "summary", fun.y= quantile, probs = .5, color = "green") +
  geom_line(stat = "summary", fun.y= quantile, probs = .9, color = "red") 
 
  
  
  
  scale_color_manual("", 
                      breaks = c("Mean", "10% quartile", "median", "90% quartile"),
                      values = c("black", "blue", "green", "red"))

#figure out how to make scale_color_manual work
```

#### What are some of your observations of the plot?
Response: Data is very dispersed around the mean


### Correlation
Notes: we'll measure the linear relationship between age and friend count using pearson's r
use ?cor.test

```{r Correlation}
cor.test(facebook$age, facebook$friend_count, method = "pearson")

#another way is by using with() and gets us the same result

with(facebook, cor.test(age, friend_count, method = "pearson"))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response: -0.027 .. .very small, so little correlation 

***

### Correlation on Subsets
Notes: want to take out users older than 70 because many people put a fake old age


```{r Correlation on Subsets}
with(facebook[facebook$age <= 70,],cor.test(age, friend_count, method = "pearson"))
```

Result:  -0.171 correlation. So taking older people out, there is a negative correlation - as age increases, friend count decreases

***

## Create Scatterplots
Notes:
www_likes_received v. likes_received

```{r}
ggplot(aes(x= www_likes_received, y = likes_received), data = facebook) + geom_point(alpha = 1/20) +
  xlim(0, quantile(facebook$www_likes_received, 0.95)) +
  ylim(0, quantile(facebook$likes_received, 0.95)) +
  geom_smooth(method = "lm", color = "red")

#I used 95th quantile to limit the axes so I get rid of outliers. Most data is located 300 www likes and 5000 likes

#added line of best fit in red 
```


What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(facebook$www_likes_received, facebook$likes_received, method = "pearson")
```

Response: 0.94, very strong correlation (obviously, bc some likes_received come from www likes, one is a subset of the other)

REMEMBER: regression assumes that the variables will be independent of each other - so high correlation between two variables messes up regression analysis

***


### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)

#load new data
data(Mitchell)
?Mitchell
```

Create plot of temperature versus Months

```{r Temp vs Month}
View(Mitchell)

ggplot(aes(x=Month, y = Temp), data = Mitchell) + geom_point() +
  geom_smooth(method = "lm", color = "red")

```

Very scattered data, line of best fit line is useless
***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot. Very low corr

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

cor.test(Mitchell$Month, Mitchell$Temp, method = "pearson")

#Result:  0.057 .. very low cor
```

***

### Making Sense of Data
Notes: Our data isn't divided into actual months, just 200 continuous months, let's add actual months that repeat, bc temp would repeat
break up Month by 12



```{r Making Sense of Data}

ggplot(aes(x=Month, y = Temp), data = Mitchell) + geom_point() +
  scale_x_discrete(breaks = seq(0, 203, 12))
```

***

### A New Perspective

What do you notice when you stretch the graph as long as possible?
Response: There is a cyclical patterns which correlation doesn't account for. Up-down bc of the seasons of the months

```{r}
ggplot(aes(x=(Month%%12),y=Temp),data=Mitchell)+ 
  geom_point() 

## %% superinposes each year's data into each other, so we can see the pattern better
```



***

### Understanding Noise: Age to Age Months
Notes: we observed that the mean friend count by age had a lot of noise.
Some noise and spikes make sense, but others are just noise on the generally smooth line bc we simply got a certain sample that is that way but isn't so for the whole population

Convert age to age with months to see all noise
```{r Understanding Noise: Age to Age Months}
ggplot(aes(x=age, y = friend_count_mean), data = facebook.fc_by_age) +
  geom_line()

```

***

### Age with Months Means

```{r Age with Months Means}
facebook$age_with_months <- facebook$age + (1- facebook$dob_month /12)
#changes age to age as decimal and more continuous
```

Programming Assignment
```{r Programming Assignment}
library("dplyr")
age_with_months <- group_by(facebook, age_with_months)

facebook.fc_by_age_month <- summarize(age_with_months,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n=n())

#created new df
```

***

### Noise in Conditional Means
plot friend_count_mean versus age_with months
```{r Noise in Conditional Means}

ggplot(aes(x=age_with_months, y = friend_count_mean), data = subset(facebook.fc_by_age_month, age_with_months <71)) +
  geom_line() 

#compare to previous one we created with ages not in months, this one has much more noise

```

***

### Smoothing Conditional Means
Notes:
COMPARE
```{r Smoothing Conditional Means}

p1 <- ggplot(aes(x=age, y = friend_count_mean), data = subset(facebook.fc_by_age, age <71)) +
  geom_line(aes(color = "red")) +
  geom_smooth()

p2<-ggplot(aes(x=age_with_months, y = friend_count_mean), data = subset(facebook.fc_by_age_month, age_with_months <71)) +
  geom_line() +
  geom_smooth()

#COMPARE and add a smooth()
library(gridExtra)
grid.arrange(p1,p2, ncol =1)
```

***

### Which Plot to Choose?
Notes:

Use many - there are tradeoffs in each one, and each plot gives you different insights

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

