Lesson 5
========================================================

### Third Qualitative Variable
Notes: We saw that women had more friends on average than men, but is it because women had a different age distribution (more younger women in the df)? Conditional on age, do women still have more friends?



```{r Third Qualitative Variable}
library(ggplot2)
ggplot(aes(x = gender, y = age), data = subset(facebook, !is.na(gender))) + geom_boxplot()
```

***

### Plotting Conditional Summaries
Notes: group by gender and plot median friend count by gender
before we had simply used conditions summary and fun.y to get the median of friend_count, here that isn't needed bc we created a new df with the median. so we just plot that


```{r Plotting Conditional Summaries}
library(dplyr)
age_gender <- group_by(facebook, age, gender)
fc_by_age_gender<- summarize(age_gender,
            mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n=n())

fc_by_age_gender <- na.omit(fc_by_age_gender)

ggplot(aes(x=age, y =median_friend_count), data = fc_by_age_gender) +
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes: How many times (x) more do women have friends than men?

***

### Wide and Long Format
Notes:
use reshape2() package and use dcast() to switch between the two formats for a dataframe
***

### Reshaping Data
Notes:
dcast(old df, vars I want to keep, ~ vars that will have their own column, value.var = "" key measurements for values)

also can use melt() - is it for df also? 

```{r}
install.packages('reshape2')
library(reshape2)
fc_wide <- dcast(fc_by_age_gender, age ~ gender, value.var = "median_friend_count")

#get a new df with each age and median friend count per gender
```


***

### Ratio Plot
Notes: Plot ratio of female to male to see how many more times do women have more friends than men and at what ages


```{r Ratio Plot}
ggplot(aes(x=age, y = (female/male)), data = fc_wide) +
  geom_line() + geom_hline(yintercept=1)
```
This showed that friend count is conditional on age

***

### Third Quantitative Variable
Notes:
tenure, how long a user has been on fb, can also affect fried count. Let's explore all three vars - friend count, age, gender, and tenure using a scatterplot and bins

```{r Third Quantitative Variable}
#create year_joined variable
facebook$year_joined <- floor(2014-facebook$tenure/365)

table(facebook$year_joined)
```

***

### Cut a Variable
Notes: Use cut() to split year_joined into separate bins 

```{r Cut a Variable}
facebook$year_joined <- as.numeric(as.character(facebook$year_joined))
facebook$year_joined.bucket <-cut(facebook$year_joined,  
          c(2004, 2009, 2011, 2012, 2014))

table(facebook$year_joined.bucket)
```

***

### Plotting it All Together
Notes:
plot friend_count v. age so each year join bucket has a line that is the median friend count

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count), 
  data = subset(facebook, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)

```

Younger users with longer tenure tend to have more friends
***

### Plot the Grand Mean
Notes:
Change above plot to mean, and plot the overall mean of friend count by ages

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count), 
  data = subset(facebook, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat= "summary", fun.y= mean, linetype = 2, aes(color = 'Grand Mean'))

```

***

### Friending Rate
Notes: Friends someone has for each day they are using fb
Create summary of friends per day of tenure, with at least one day of tenure

```{r Friending Rate}
with(subset(facebook, tenure >=1), summary(friend_count/tenure))
```

***

### Friendships Initiated
graph friendships initiated per day versus tenure

```{r Friendships Initiated}
ggplot(aes(x=tenure, y = friendships_initiated/tenure),
       data = subset(facebook, tenure >= 1)) +
       geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)

###add a smoother to remove noise by changing geom_line to geom_smooth and remove summary and mean
ggplot(aes(x=tenure, y = friendships_initiated/tenure),
       data = subset(facebook, tenure >= 1)) +
       geom_smooth(aes(color = year_joined.bucket))

```

users with more tenure initiate less friendships

***

### Bias-Variance Tradeoff Revisited
Notes:

Following code changes bin of x to reduce noise
variance:  having a lot of noise
bias:  reducing variance creates a bias in the data
```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***


### Introducing the Yogurt Data Set
Notes:
purchases of yogurt
***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yogurt <-read.csv("/Users/danilitovsky/Dropbox/dataScience/R/yogurt.csv")

#convert id variable to a factor
yogurt$id <- factor(yogurt$i
str(yogurt)

ggplot(aes(x=price), data = yogurt) + geom_histogram()

table(yogurt$price)
```

***

### Number of Purchases
Notes: Create a new variable combining all the yogurt flavors purchased each time 

```{r Number of Purchases}
yogurt$all_purchases <-yogurt$strawberry + yogurt$blueberry + yogurt$pina.colada + yogurt$plain + yogurt$mixed.berry

ggplot(aes(x=all_purchases), data = yogurt) + geom_histogram()

#most buy 1-2 yogurts at a time

```

***

### Prices over Time
Notes: create scatter of price vs. time, asses for overplotting

```{r Prices over Time}
ggplot(aes(x=time, y = price), data= yogurt) +
  geom_jitter(alpha = 1/4, shape =21)

#most common prices went up over time, points scattered below may be due to using coupons
```

***

### Sampling Observations
Notes: Let's look at a subsample of households

***

### Looking at Samples of Households
set.seed() is like nrom() in that it produces random numbers but you can save the numbers you get

```{r Looking at Sample of Households}
#set the seed for reproducible results. 
set.seed(4230)
sample.ids <- sample(levels(yogurt$id), 16)
#gets me 16 sample households (each house has an id) with the same seed number, use levels bc I want different households

sample.ids #prints sample ids out

#plot each purchase occasion for each of the 16 sample households

ggplot(aes(x=time, y= price), data = subset(yogurt, id %in% sample.ids)) +
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size = all_purchases), pch= 1)
  
# x %in% y returns a logical boolean vector stating if x appears in y, so we only get the households from the dataset with the sample.ids, it creates a loop over ids

#size = all_purchases will alter size depending on how many yogurts a household purchased

###replicate w different seed
set.seed(333)
sample.ids <- sample(levels(yogurt$id), 16)

sample.ids 

ggplot(aes(x=time, y= price), data = subset(yogurt, id %in% sample.ids)) +
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size = all_purchases), pch= 1)

```

Graph:  shows how over time, which households consistently buy yogurt or not, and for most households, the price goes up over time, and how much they buy at once (size)
***


### Many Variables
Notes: Use multiple plots in scatterplot matrix

***

### Scatterplot Matrix
Notes: with GGally package

```{r}
install.packages("GGally")
library(GGally)
theme_set(theme_minimal(20))

##set the seed for reproducible results
set_seed(1836)
facebook_subset <- facebook[, c(2:15)] #subset only certain variables (not categorical vars)

ggpairs(facebook_subset[sample.int(nrow(facebook_subset), 1000),])
####  WARNING :   ABOVE CODE TAKES 1 HR TO LOAD !!!

```


***

### Even More Variables
Notes:

***

### Heat Maps
Notes: genomic data to create a heat map
use red to show when a gene is overexpressed and blue when it is underexpressed


```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
##change data to long format
##only looking at first 200 of over 6000 genes
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```
***
***

####more practice
Create new var of proportion of friendships that a user initiated
# Create a line graph of the proportion of
# friendships initiated ('prop_initiated') vs.
# tenure and color the line segment by
# year_joined.bucket. Use mean to remove overplotting
```{r}
library(gridExtra)
facebook2$prop_initiated <- facebook2$friendships_initiated / facebook2$friend_count

p1 <- ggplot(aes(x=tenure, y = prop_initiated), data= facebook2) +
  geom_line(aes(color=year_joined.bucket), stat = "summary", fun.y = mean)

###with smoother and big bin size
p2 <-ggplot(aes(x=tenure, y = prop_initiated, binwidth = 2000), data= facebook2) +
  geom_smooth(aes(color=year_joined.bucket), stat = "summary", fun.y = mean)

grid.arrange(p1, p2)

##not much change with smoothers

###what's the avg proportion of friendships initiated by the 2012-2014 group?

by(facebook2$prop_initiated, facebook2$year_joined.bucket, summary)
```

Conclusion:  more recent users have a higher proportion of friend initiation (maybe bc there are more people using fb when they join), but users from the beginning of facebook have a huge proportion variance, some initiate and some don't.



Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

